### 2020-08-14 paper.Rmd
Paper update

---
title: "CES Package Writeup"
author: "Paul A. Hodgetts"
date: "13/08/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
In `CESR`, we provide a means of accessing the Canadian Election Study survey datasets in one place that does not require storing data files nor the reading of a file from a working directory. Instead, the `CESR` temporarily downloads a compressed .zip file of a requested CES survey dataset, unzips the file and reads the dataset into R as a data object, and then removes the downloaded and unzipped files from the computer. Additionally, the `CESR` provides a subset of the 2019 CES online survey that can can be accessed through its own function call.

The `CESR` is important because it makes working with Canadian Election Study survey datasets easier. By circumventing the need the find, download, and read in a dataset, `CESR` removes the requirement of setting a working directory thereby making the process of setting up a data file for use in R much easier. This makes working between computers much easier, meaning that R projects can easily be shared between team members without the concern of a file being properly read or code not working due to different working directories. Additionally, by creating a subset of the CES 2019 online survey through a built-in function call, the `CESR` provides educators with an important tool that can be used to aid in the teaching of the statistical exploration of survey data.

Our package is functionally complimentary to working in R Studio by minimizing the number of steps required to load in a dataset and removing the need of storing data file. Furthermore, the `CESR` package complements the work being done in the R community by building on the work being done through other packages that pull in survey data such as `dataverse` and `cancensus`.

# Key functions
The main function in `CESR` is.... This takes the following arguments... An example use case for this is...

# Supporting functions
Supporting functions in `CESR` include...

# Vignette

## Getting the 2019 phone CES

## Something else?

# Next steps and cautions